A Temporal Logic Based Framework for Intrusion Detection

### 2023-08-31 Paper/main.tex
update

@@

Assume that we want to state the following property: "\textit{Whenever there is a login by any user x, then eventually the user x logs out}". In LTL we can write the following formula: 
\begin{align*}
    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =\ userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
\end{align*}
In this formula we use the operator \underline{let} \_ \underline{in} \_ to bind the value of $userId$ in the current event to the local variable $k$ whenever $\text{(action = login)}$ in the current event; we then impose the condition that the value of $userId$ in some event in future must be same as the user id bound to k and that the action of the event must be logout.
In EAGLE we can express the same property with the following rules:
\begin{align*}
    & \underline{\text{min}}\ \text{EvLogout}(\underline{\text{string}}\ k) = (\text{action = logout}\land userId = k) \lor \bigcirc \text{EvLogout}(k) \\
    & \underline{\text{mon}}\ M_2 = ((\text{action = login})\rightarrow \text{EvLogout}(userId)) 
\end{align*}
As a result, rules in EAGLE give us the power to create specific temporal operators as well as to bind and modify data. This property turns out to be crucial for succinctly expressing executions of attack-safe systems. To make the paper self-contained, we go back to the syntax and semantics of EAGLE.
\subsection{Syntax and Semantics}
A specification $S$ is made up of an observer part $O$ and a declaration part $D$. $O$ is made up of zero or more monitor definitions $M$, which define what will be watched, while $D$ is made up of zero or more rule definitions $R$. The names of rules and monitors are $(N)$.
\begin{align*}
    S &::= D \, O \\
    D &::= R^* \\
    O &::= M^* \\
    R &::= \{\underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
    M &::= \underline{\text{mon}}\ N = F \\
    T &::= \underline{\text{Form}}\ |\ \text{primitive type} \\
    F &::= \text{expression}\ |\ \underline{\text{true}}\ |\ \underline{\text{false}}\ |\ \neg F\ |\ F_1 \land F_2\ |\ F_1 \lor F_2\ |\\
            &F_1 \rightarrow F_2\ |\ \bigcirc F\ |\ \odot F\ |\ F_1 \cdot F_2\ |\ N(F_1, \ldots, F_n)\ |\ x_i
\end{align*}
A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$. While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); as we will see, the difference becomes important only when we are evaluating the formula in specific instants. Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored.

>>>>

%% INTRODURRE VELOCE ESEMPIO E STRUTTURA DEL PROGRAMMA

@@

Assume that we want to state the following property: "\textit{Whenever there is a login by any user x, then eventually the user x logs out}". In LTL we can write the following formula: 
\begin{align*}
    & \square{(\text{(action = login)}\rightarrow \underline{\text{let}}\ \text{k} =\ userId\ \underline{\text{in}}\ \lozenge(\text{action = logout}\land userId = k))}
\end{align*}
In this formula we use the operator \underline{let} \_ \underline{in} \_ to bind the value of $userId$ in the current event to the local variable $k$ whenever $\text{(action = login)}$ in the current event; we then impose the condition that the value of $userId$ in some event in future must be same as the user id bound to k and that the action of the event must be logout.
In EAGLE we can express the same property with the following rules:
\begin{align*}
    & \underline{\text{min}}\ \text{EvLogout}(\underline{\text{string}}\ k) = (\text{action = logout}\land userId = k) \lor \bigcirc \text{EvLogout}(k) \\
    & \underline{\text{mon}}\ M_2 = ((\text{action = login})\rightarrow \text{EvLogout}(userId)) 
\end{align*}
As a result, rules in EAGLE give us the power to create specific temporal operators as well as to bind and modify data. This property turns out to be crucial for succinctly expressing executions of attack-safe systems. To make the paper self-contained, we go back to the syntax and semantics of EAGLE.
\subsection{Syntax and Semantics}
A specification $S$ is made up of an observer part $O$ and a declaration part $D$. $O$ is made up of zero or more monitor definitions $M$, which define what will be watched, while $D$ is made up of zero or more rule definitions $R$. The names of rules and monitors are $(N)$.
\begin{align*}
    S &::= D \, O \\
    D &::= R^* \\
    O &::= M^* \\
    R &::= \{\underline{\text{max}}\ |\ \underline{\text{min}} \}\ N(T_1\ x_1, \ldots, T_n\ x_n) = F \\
    M &::= \underline{\text{mon}}\ N = F \\
    T &::= \underline{\text{Form}}\ |\ \text{primitive type} \\
    F &::= \text{expression}\ |\ \underline{\text{true}}\ |\ \underline{\text{false}}\ |\ \neg F\ |\ F_1 \land F_2\ |\ F_1 \lor F_2\ |\\
            &F_1 \rightarrow F_2\ |\ \bigcirc F\ |\ \odot F\ |\ F_1 \cdot F_2\ |\ N(F_1, \ldots, F_n)\ |\ x_i
\end{align*}
The types of parameters include primitive types like \underline{int}, \underline{long}, \underline{float}, etc. or formulas of type \underline{Form}.
A term indicating whether the interpretation is maximal ($\underline{\text{max}}$) or minimal ($\underline{\text{min}}$) comes before a rule specification $R$. While minimal rules establish \textit{liveness properties} (something good eventually happens), maximal rules define \textit{safety properties} (nothing bad ever happens); as we will see, the difference becomes important only when we are evaluating the formula in specific instants. Recursive definitions of rules are permitted as long as they are tightly guarded by a temporal operator. Finally,the definitions starting with keyword $\underline{\text{mon}}$ specifies the EAGLE formulas to be monitored and cannot have a recursive definition. We shall see that monitored rules will evolve as new events appear.
